Measurement of the \ce{^{64}Zn},\ce{^{47}Ti}(n,p) Cross Sections using a DD Neutron Generator for Medical Isotope Studies

### 2018-12-11 First_np_Paper/np_body_text.tex
Pushing minor edits made to the HFNG (n,p) paper during my dissertation editing

For the  \ce{^{47}Sc} production experiments, the foils were counted simultaneously using a planar LEPS detector.
For the \ce{^{64}Cu} production experiments, the Indium foil was first counted separately using an 80\% HPGe detector, to capture the short-lived Indium activities.
This is due to the fact that the contaminant \ce{^{115}In}(n,$\gamma$) reaction results in the production of \ce{^{116m}In} which has a 54 min half-life and results in the production of 1097 keV (58.5\% branching), 1293 keV (84.8\% branching) and 2112 keV (15.09\% branching) gamma-rays that in turn produce a significant number of 511 keV gammas from pair-production followed by annihilation \cite{Blachot2010}.
The foils were counted together again after approximately 4 h of separate collection, to allow for nearly all of the produced \ce{^{116}In} to decay.
% Example spectra for each production pathway can be seen in Figures \autoref{fig:spectra_a} and \autoref{fig:spectra_b}.
Example spectra for each production pathway can be seen in Figures \autoref{fig:main_spectra}a and \autoref{fig:main_spectra}b.

@@

An optimal design for the neutron generator would also allow for the possibility of reflecting fast neutrons back onto the target to maximize their utilization for radionuclide production.
 This sort of \enquote{flux trap} has been used for the production of radionuclides in reactors, but has not to date been optimized for use with fast neutrons  at DD and/or DT neutron sources.
 The HFNG, with its self-loading target and \enquote{flux trap} geometry, has many features that make it well-suited for such isotope production purpose.
 Switching to DT operation would dramatically increase the flux as well as the production cross section, since (n,p) tends to be significantly larger at 14 MeV.
 However, the higher neutron energy would also open the (n,pn) channels.
 In the case of \ce{^{47}Sc}, this would lead to the presence of \ce{^{46}Sc} ($t_{1/2}$ = 83.79 d \cite{Wu2000}) in the sample, which might pose some concerns for medical applications.
 However, this is not an issue for \ce{^{64}Cu} since the (n,pn) channel leads to the production of stable \ce{^{63}Cu}.

Assuming a neutron flux of  $\sci{1.3}{7}$ neutrons/cm$^2$s on the target, masses of 0.533 g of natural zinc and 0.337 g of natural titanium, and cross sections of 47.5 mb for \ce{^{64}Zn}(n,p)\ce{^{64}Cu}   and 26.26  mb for  \ce{^{47}Ti}(n,p)\ce{^{47}Sc}, theoretical saturation activities for current operation at the time of this work are estimated to be 1.5 kBq of \ce{^{64}Cu} and 0.11 kBq of \ce{^{47}Sc}.
This falls short of the mCi (37 MBq) level required for commercial application by a factor of 3-4 orders of magnitude, but with the operation of the second deuterium ion source, increased current, and fast neutron reflection, this goal may  well be within reach. 
By increasing the activation target thickness to 1 cm (a factor of 10), switching to DT operation (a factor of 80), increasing current and running the second ion source (a factor of 60), and relying upon the higher (n,p) cross section at DT energies (a factor of approximately 3), we believe saturation activities of approximately 6 mCi of \ce{^{64}Cu} and 0.5 mCi of \ce{^{47}Sc} can be achieved.  
The activities produced at the end of irradiation averaged 453.8 Bq of \ce{^{64}Cu}, and 31.6 Bq of \ce{^{47}Sc}.
Assuming a conservative neutron source output of $10^8$ neutrons / second, we can estimate that, in present operation, the HFNG has an average $\eta_{64Cu} \approx \sci{3.0}{-5}$ for \ce{^{64}Cu} and $\eta_{47Sc} \approx \sci{1.1}{-5}$ for  \ce{^{47}Sc}. 
This falls approximately 4 orders of magnitude short of the $\eta_x \approx 0.37$ needed for mCi-scale production.
A factor of 10 in $\eta_x$ could easily be gained through use of targets  1-cm in thickness without worry of contaminating reaction channels opening up, but $\eta_x$ gains beyond this will require modification of operation conditions.

@@

 In addition, the   production of the \ce{^{116}In}  via the \ce{^{115}In}(n,$\gamma$) reaction was close to the value one would expect given an effective incident neutron energy of 2.45 MeV.  
 While this is not consistent with the average neutron energy at the target location (2.76$^{+0.01}_{-0.02}$ MeV), the fact that it was close indicates the paucity of thermal neutrons in this central location.  
 This in turn highlights the usefulness of such compact DD-neutron sources  for producing \enquote{clean} activities via the (n,p) channel.  
The use of DD neutron generators can be an efficient method for the measurement of low-energy (n,p) reaction channels, as well as a relative method used to normalize measurements at higher neutron energies.
 In addition to improving the value of these measurements for nuclear reaction evaluation, our results highlight the potential use of compact neutron generators for the production of radionuclides locally for medical applications.

>>>>

For the  \ce{^{47}Sc} production experiments, the foils were counted simultaneously using a planar LEPS detector.
For the \ce{^{64}Cu} production experiments, the Indium foil was first counted separately using an 80\% HPGe detector, to capture the short-lived Indium activities.
This is due to the fact that the contaminant \ce{^{115}In}(n,$\gamma$) reaction results in the production of \ce{^{116m}In} which has a 54 min half-life and results in the production of 1097 keV (58.5\% branching), 1293 keV (84.8\% branching) and 2112 keV (15.09\% branching) gamma-rays that, in turn, produce a significant number of 511 keV gammas from pair-production followed by annihilation \cite{Blachot2010}.
The foils were counted together again after approximately 4 h of separate collection, to allow for nearly all of the produced \ce{^{116}In} to decay.
% Example spectra for each production pathway can be seen in Figures \autoref{fig:spectra_a} and \autoref{fig:spectra_b}.
Example spectra for each production pathway can be seen in  
% Figures 
\autoref{fig:main_spectra}a and \autoref{fig:main_spectra}b.

@@

An optimal design for the neutron generator would also allow for the possibility of reflecting fast neutrons back onto the target to maximize their utilization for radionuclide production.
 This sort of \enquote{flux trap} has been used for the production of radionuclides in reactors, but has not, to date, been optimized for use with fast neutrons  at DD and/or DT neutron sources.
 The HFNG, with its self-loading target and \enquote{flux trap} geometry, has many features that make it well-suited for such isotope production purpose.
 Switching to DT operation would dramatically increase the flux as well as the production cross section, since (n,p) tends to be significantly larger at 14 MeV.
 However, the higher neutron energy would also open the (n,pn) channels.
 In the case of \ce{^{47}Sc}, this would lead to the presence of \ce{^{46}Sc} ($t_{1/2}$ = 83.79 d \cite{Wu2000}) in the sample, which might pose some concerns for medical applications.
 However, this is not an issue for \ce{^{64}Cu} since the (n,pn) channel leads to the production of stable \ce{^{63}Cu}.

Assuming a neutron flux of  $\sci{1.3}{7}$ neutrons/cm$^2$s on the target, masses of 0.533 g of natural zinc and 0.337 g of natural titanium, and cross sections of 47.5 mb for \ce{^{64}Zn}(n,p)\ce{^{64}Cu}   and 26.26  mb for  \ce{^{47}Ti}(n,p)\ce{^{47}Sc}, theoretical saturation activities for current operation at the time of this work are estimated to be 1.5 kBq of \ce{^{64}Cu} and 0.11 kBq of \ce{^{47}Sc}.
This falls short of the mCi (37 MBq) level required for commercial application by a factor of 3-4 orders of magnitude, but,  with the operation of the second deuterium ion source, increased current, and fast neutron reflection, this goal may  well be within reach. 
By increasing the activation target thickness to 1 cm (a factor of 10), switching to DT operation (a factor of 80), increasing current and running the second ion source (a factor of 60), and relying upon the higher (n,p) cross section at DT energies (a factor of approximately 3), we believe saturation activities of approximately 6 mCi of \ce{^{64}Cu} and 0.5 mCi of \ce{^{47}Sc} can be achieved.  
The activities produced at the end of irradiation averaged 453.8 Bq of \ce{^{64}Cu}, and 31.6 Bq of \ce{^{47}Sc}.
Assuming a conservative neutron source output of $10^8$ neutrons / second, we can estimate that, in present operation, the HFNG has an average $\eta_{64Cu} \approx \sci{3.0}{-5}$ for \ce{^{64}Cu} and $\eta_{47Sc} \approx \sci{1.1}{-5}$ for  \ce{^{47}Sc}. 
This falls approximately 4 orders of magnitude short of the $\eta_x \approx 0.37$ needed for mCi-scale production.
A factor of 10 in $\eta_x$ could easily be gained through use of targets  1-cm in thickness without worry of contaminating reaction channels opening up, but $\eta_x$ gains beyond this will require modification of operation conditions.

@@

 In addition, the   production of the \ce{^{116}In}  via the \ce{^{115}In}(n,$\gamma$) reaction was close to the value one would expect given an effective incident neutron energy of 2.45 MeV.  
 While this is not consistent with the average neutron energy at the target location (2.76$^{+0.01}_{-0.02}$ MeV), the fact that it was close indicates the paucity of thermal neutrons in this central location.  
 This, in turn, highlights the usefulness of such compact DD-neutron sources  for producing \enquote{clean} activities via the (n,p) channel.  
The use of DD neutron generators can be an efficient method for the measurement of low-energy (n,p) reaction channels, as well as a relative method used to normalize measurements at higher neutron energies.
 In addition to improving the value of these measurements for nuclear reaction evaluation, our results highlight the potential use of compact neutron generators for the production of radionuclides locally for medical applications.
